Nail: A practical interface generator for data formats

### 2014-04-29 paper/osdi14/design.tex
Misc paper changes

Nail grammars describe not only a

To integrate the data format and the internal representation, Nail
provides a rich set of combinators that not
only describe the grammar of the external protocol, but also induce an internal
model.

A central design decision of Nail is that there is a \emph{semantic bijection} between
the model exposed to the programmer and the byte-level input and output, up to
syntactic equivalence for unambiguous grammars. More precisely, the parser is
the generator's inverse, so parsing the generator's output will yield the
generator's input, but generating the parsers output does not necessarily yield
the parsers input. To understand why this makes sense, consider a grammar for a
text language that tolerates white space, or a binary protocol that tolerates
arbitrarily long padding.\footnote{Say, the physical layer of most communication
  protocols is a possibly infinite sequence of symbols that are syntactically
  nil followed by a pre-determined synchronization sequence and the actual
  contents of the transmission.} In that case, the program semantics should be
independent of the number of padding elements in the input, and Nail therefore
does not expose that information to the programmer. We call such discarded
fields \emph{constants}. Currently, Nail always makes a default choice when
there are multiple options to express a constant, however Nail could be extended
to allow a grammar-specific plug-in to make these choices, say for faster
alignment, streaming applications when data is not ready, or visual appearance in
human-readable protocols. Similarly, Nail does not preserve the layout of
objects referred to by their offsets. If the grammar contains no constants and
offset fields, there is a proper isomorphism between model and protocol.

Nail evaluates the combinators and produces:

@@

\noindent
In the rest of this section, we present Nail's combinators.

\begin{figure}[tb]
\begin{tabular}{ll}
\toprule
\bf Syntax & \bf Semantics \\
\midrule
\texttt{int32} & 32-bit signed integer \\
\multirow{2}{*}{\texttt{uint4}}
  & 4-bit unsigned integer, \\
  & returned as an 8-bit value \\
\multirow{2}{*}{\texttt{uint8 | 1..3}}
  & 8-bit unsigned integer, \\
  & $1\leq x \leq 3$ \\
\multirow{2}{*}{\texttt{uint16 | ..512}}
  & unsigned 16-bit integer, \\
  & $x \leq 512$ \\
\multirow{2}{*}{\texttt{int32 | [1,255,512]}}
  & signed 32 bit integer, \\
  & $x \in \{ 1, 255, 512 \}$ \\
\bottomrule
\end{tabular}
\caption{Example Nail grammars for integer values.}
\label{fig:range}
\end{figure}

>>>>

Nail grammars describe both the \emph{external format} and an \emph{internal representation} of a
protocol. Typical grammar languages describe only the former and leave it up to the programmer to
construct an internal representation, typically in the form of an abstract syntax tree.

Our grammars are based on Parser Expression Grammars, describing

Nail produces the following from a single, descriptive grammmar:

@@

\noindent
In the rest of this section, we present the design of the Nail language.

\subsection{Basics}
Nail grammars contain parsers
\begin{figure}[tb]
\begin{tabular}{ll}
\toprule
\bf Syntax & \bf Semantics \\
\midrule
\texttt{int32} & 32-bit signed integer \\
\multirow{2}{*}{\texttt{uint4}}
  & 4-bit unsigned integer, \\
  & returned as an 8-bit value \\
\multirow{2}{*}{\texttt{uint8 | 1..3}}
  & 8-bit unsigned integer, \\
  & $1\leq x \leq 3$ \\
\multirow{2}{*}{\texttt{uint16 | ..512}}
  & unsigned 16-bit integer, \\
  & $x \leq 512$ \\
\multirow{2}{*}{\texttt{int32 | [1,255,512]}}
  & signed 32 bit integer, \\
  & $x \in \{ 1, 255, 512 \}$ \\
\bottomrule
\end{tabular}
\caption{Example Nail grammars for integer values.}
\label{fig:range}
\end{figure}

### 2014-04-29 paper/osdi14/eval.tex
Misc paper changes

\section{Evaluation}
\label{s:eval}
\input{fig-ethernet}
\input{fig-dnsfull}
\input{fig-zip}
In our preliminary evaluation of Nail, we try to answer four questions:

\begin{itemize}
\item Can Nail grammars support real-world data formats?
\item How much programmer effort is required to build an
      application that uses Nail for data input and output?
\item Do applications handling input and output with Nail suffer from less security vulnerabilities?
\item How fast are systems using Nail parsers as opposed to hand-written input handling? 
\end{itemize}

\subsection{Data formats.}
\label{s:eval-formats}
To answer the first question, we implemented grammars for four protocols with particularly
challenging features and show that Nail is a good fit for these data formats. In
\S\ref{s:eval-effort} we will demonstrate that Nail is easier to use than hand-written parsers and
more convenient than state-of-the-art parser generators for formats that they can handle.

In particular, we show extracts from grammars for the following protocols: 
\begin{itemize}
\item A subset of Ethernet network packets, containing ARP, ICMP and UDP/IP packets. See Figure~\ref{fig:ethernet-extract}.
\item DNS packets as described in RFC 1035, featuring label compression. See Figure~\ref{fig:dns-full}.
\item ZIP files, with compression, check-sums and a complicated procedure to identify the header.
  See Figure~\ref{fig:zip-extract}.
\end{itemize}

one for  DNS queries and respones, as per RFC1035 (shown in, one for the ZIP
archive format and a grammar for Ethernet frames containing ARP, ICMP and UDP packets. We also
parse Nail grammars with Nail, demonstrating that our system can also handle text-based protocols
and languages.

@@

The first application we built with Nail, is a toy DNS server inspired by the example shipped with
Hammers DNS server consists of 683 lines of C, excluding the Hammer framework itself, that responds to
any valid DNS query with a CNAME record to the domain ``spargelze.it''.
Most of this code is taken up with custom validators, semantic actions,
and data structure definitions, with only 52 lines of code defining the
grammar with Hammer's combinators.

@@

Our simple DNS server, which parses a zone file, listens to incoming DNS requests, parses them and
generates appropriate responses with this Nail grammar, is implemented in 149 lines of C.

 The same grammar is used, together with 98 lines of C,
to implement a functional toy clone of the \texttt{host} command-line
tool.

It is hard to compare the programming effort required to implement
our toy DNS server to that of a real world DNS server, since
we have less , in particular the additional hint records that real-world DNS servers send and a
complete lack of configuration options.
However, the closest in functionality and intent is Dan Bernstein's
djbdns,\footnote{\url{http://cr.yp.to/djbdns.html}} which aims to be
a minimalist, highly secure DNS server. The latest release of djbdns,
including various support tools, is about 10,000 lines of C as measured by
\texttt{sloccount}. \XXX[Show that we only count the DNS server parts, not the rest. List some
features we don't have]

@@

Another particularly tricky file format is the ZIP compressed archive format\cite{pkzip}. The ZIP
format has been associated with many vulnerabilities, and the PROTOS Genome project found numerous
security vulnerabilities in most implementations of ZIP and other archive formats. Furthermore,
Android security has been completely by-passed multiple times through parsing differentials in
handling ZIP files\cite{saurik-masterkey}. \XXX[mention talk by Julia Wolf about zip].

ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an \emph{end-of-directory
header}. This header contains a variable length comment, so it has to be located by scanning
backwards from the end of the file until a magic number and a valid length field is found. Many ZIP
implementations disagree on how to find this header in confusing situations, such as when the
comment contains the valid magic bytes\cite{wolf-berlinsides-zip}.
This end-of-directory header contains the offset and size of the \emph{ZIP directory}, which is an
array of one \emph{directory entry header} per file in the archive.
Each entry stores file meta data, such as file name, compressed and uncompressed size and a checksum,
in addition to the offset of a \emph{local file header}. The local file header duplicates most
information from the directory entry header and the compressed file contents follow it immediately.

\XXX[condense]

We wrote a grammar for ZIP, which consists of 92 lines of Nail and 78 lines of C implementing two
stream transforms, one for the DEFLATE compression algorithm with the help of the zlib library and
one for finding the end-of-directory header. This grammar avoids the two pitfalls mentioned above:
Nail forces the programmer to specify an explicit procedure for finding the end-of-directory record
and we use dependent fields to enforce that the duplicated information is consistent.

@@

\paragraph{Network stack.}
Write up this stack.
\subsection{Security.} 
Fuzz DNS. Fuzz ZIP. Design argument that security is hard to show.

\subsection{Performance.}

Compare DNS requests per second of BIND, DJBDNS and NailDNS on a small zonefile.
Compare extraction performance for ZIP.

>>>>

\section{Evaluation}
\label{s:eval}
\input{fig-ethernet}
\input{fig-dnsfull}
\input{fig-zip}
In our evaluation of Nail, we try to answer four questions:

\begin{itemize}
\item Can Nail grammars support real-world data formats and are Nail's concepts critical to handling
  them?
\item How much programmer effort is required to build an
      application that uses Nail for data input and output?
\item Do applications handling input and output with Nail suffer from less security vulnerabilities?
\item How fast are systems using Nail parsers as opposed to hand-written input handling? 
\end{itemize}

\subsection{Data formats.}
\label{s:eval-formats}
To answer the first question, we implemented grammars for four protocols with particularly
challenging features and show that Nail is a good fit for these data formats and some aspects of
these formats would be very hard to parse with other parser generators.

In particular, we show extracts from grammars for the following protocols: 
\begin{itemize}
\item DNS messages.
\item Ethernet packets containing ARP, ICMP and UDP.
\item ZIP files.
  See Figure~\ref{fig:zip-extract}.
\end{itemize}

\paragraph{DNS.}
Domain Name System(DNS) packets are described in RFC 1035 and our grammar almost directly
corresponds to the diagrams in Section 4 of that document. Each DNS packet consists of a header, a
set of question records and a set of answer records. Domain names in both queries and answers are
encoded as a sequence of Pascal-style strings. Each label is preceded by a length field, and the
domain name is terminated by a $0$ octet. The client's query packet contains one question
record to the server for each DNS look-up it wants to perform. In addition to the appropriate
response records, the server's reply contains a duplicate of the client's question section, which
allows stateless DNS clients and caches.

To reduce the size overhead of including each domain name twice in a DNS reply, once in the
question section, at least once in the response record, DNS includes a label compression scheme. If
a domain name suffix is repeated, instead of repeating that suffix, the DNS packet may write a
two-bit marker sequence followed by a 14-bit offset into the packet where that suffix was previously
encoded.

Processing these encodings in existing tools, such as Bison or Hammer, would at best be very
awkward, because some ad-hoc trick would have to be used to re-position the parsers input stream.
Keeping track of the position all recognised labels would not be enough, as the offset field may
refer to any byte within the packet, not just the beginning of labels. For this reason, the DNS
server used as a tutorial example for Hammer does not support compression.

Nail handles compression with a stream transform. When parsing, this transform decompresses the DNS
label stream by following the offset pointers. When generating, this transform receives the current
suffix as an input and can scan the packet so far for previous occurrences.

\paragraph{Ethernet}
\XXX[TODO]
\paragraph{ZIP Files.}

Another particularly tricky file format is the ZIP compressed archive format\cite{pkzip}.
ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an \emph{end-of-directory
header}. This header contains a variable length comment, so it has to be located by scanning
backwards from the end of the file until a magic number and a valid length field is found. Many ZIP
implementations disagree on how to find this header in confusing situations, such as when the
comment contains the valid magic bytes\cite{wolf-berlinsides-zip}.
This end-of-directory header contains the offset and size of the \emph{ZIP directory}, which is an
array of one \emph{directory entry header} per file in the archive.
Each entry stores file meta data, such as file name, compressed and uncompressed size and a checksum,
in addition to the offset of a \emph{local file header}. The local file header duplicates most
information from the directory entry header and the compressed file contents follow it immediately.

Furthermore, Nail transparently handles ZIP's data compression and checksums with stream
transformations. The implementation of one of these transforms are shown in
Figure~\ref{fig:zip-compress}. For formats such as Office documents or Android applications that are based on ZIP
containers, our ZIP grammar could be modified by including a more concrete grammar for file
contents, which currently are just an array of bytes.

With existing parser languages, the grammar can not guarantee correctness of checksums, length
fields or offsets, which places additional verification burden on the programmer. Furthermore,
existing parser generators cannot decompress data as they are parsing. This is merely inconvenient
for writing a grammar for plain ZIP files, but this prevents the grammar from validating the data
within a ZIP file.

@@

@@

With this grammar, we built a simple authoritative DNS server, which parses a zone file, listens to incoming DNS
requests, parses them and generates appropriate responses, is implemented in 183 lines of C. The
same grammar is used, together with 98 lines of C, to implement a minimal clone of the
\texttt{host} command-line tool. Most of this code consists of a custom hashtable and system
interface code, such as listening to sockets and reading the zonefile.

The Hammer parser framework\cite{hammer-dns} ships with a toy DNS server that responds to
any valid DNS query with a CNAME record to the domain ``spargelze.it''. 
The server consists of 683 lines of C, mostly custom validators, semantic actions,
and data structure definitions, with only 52 lines of code defining the
grammar with Hammer's combinators.

It is hard to compare the programming effort required to implement
our DNS server to that of a real world DNS server, since we implement less functionality.
 In particular, we do not send the additional glue records real-world DNS servers send and
we did not implement any configuration options.

However, the closest in functionality and intent is Dan Bernstein's
djbdns,\footnote{\url{http://cr.yp.to/djbdns.html}} which aims to be
a minimalist, highly secure DNS server. The latest release of djbdns,
including various support tools, is about 10,000 lines of C as measured by
\texttt{sloccount}. \XXX[Show that we only count the DNS server parts, not the rest. List some
features we don't have]

@@

We wrote a grammar for ZIP, which consists of 92 lines of Nail and 78 lines of C implementing two
stream transforms, one for the DEFLATE compression algorithm with the help of the zlib library and
one for finding the end-of-directory header.

@@

\paragraph{Network stack.}
\XXX[Write this up]
\subsection{Security.} 
 \XXX[The ZIP format has been associated with many vulnerabilities, and the PROTOS Genome project found numerous
security vulnerabilities in most implementations of ZIP and other archive formats. Furthermore,
Android security has been completely by-passed multiple times through parsing differentials in
handling ZIP files\cite{saurik-masterkey}. mention talk by Julia Wolf about zip]. 
Fuzz DNS. Fuzz ZIP. Design argument that security is hard to show.

\subsection{Performance.}
\input{fig-perf-dns}
To estimate the performance impact of using Nail grammars as opposed to hand-written grammars, we
benchmarked our DNS server against BIND \XXX[and our network stack against \texttt{lwip-tap}, a userspace
router implemented with the LWIP network stack\cite{lwip}].

\paragraph{DNS}
We compare the performance of our DNS server to the release 9.9.5 of ISC BIND 9\cite{bind8}, one of
the oldest and most popular DNS daemons, on a load that resembles the authoritative name server at a
hosting company. First, we generated domain names consisting of one or
two labels randomly selected from an English dictionary and one label that is one of three popular
Top Level Domains: ``com'', ``net'' or ``org''. Second, we randomly selected 90\% of these domains and
created a zone file that mapped these domain names to the local IP address.

Finally, we used the \texttt{queryperf} tool provided with BIND to query each domain between zero
and three times on a local instance of BIND or NailDNS. Both DNS servers were operating on a single, 
core  of an Intel i7-3610QM with 12GB of RAM on an idle system, and the benchmark tool kept at most 20
queries outstanding at once. The queryperf tool was configured to
repeat the same, randomised sequence of queries for one minute and the average throughput and
response latency were measured.  We repeated each test seven times  with 50.000 and one million
domain names, restarting each daemon in between. We also performed one initial dry run to warm the
file system cache for the zone file.
\XXX[Add 1 million results.They are largely the same]

Surprisingly, our simple Nail server outperforms the established BIND server that has seen decades
of optimisation by a factor of 3. \XXX[Why the hell?] This demonstrates that even though Nail was
not developed focusing on performance, applications built on Nail can compete with real-world
hand-written parsers.

### 2014-04-29 paper/osdi14/intro.tex
Misc paper changes

Code that handles untrusted inputs, such as processing network
data or parsing a file, is error-prone and is often exploited by
attackers.  This is in part because attackers have precise control
over the inputs to that code, and can craft inputs that trigger
subtle corner cases in input processing.  For example, the libpng
library has had 24 remotely exploitable vulnerabilities from 2007 to
2013,\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html}}
and Adobe's PDF and Flash viewers have been notoriously
plagued by input processing vulnerabilities.  Even relatively
simple formats, such as those used by the zlib compression
library, have had input processing vulnerabilities in the
past.\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}

@@

Third, data formats, especially binary formats such as PNG or PDF, can have \emph{structural dependencies}, such as offset, length, and checksum fields
that are hard to express in state-of-the-art grammar languages. Often the programmer is required to
manually write control code, such as re-positioning the parser's input stream, looping over the
invocation of a sub-parser, or computing a
checksum over raw input bytes.  Besides leaving much room for errors with offset arithmetic, such
code is usually not reusable when generating output.

@@

Second, this well-defined internal representation allows Nail to invert the
parser and generate output. However, Nail allows \emph{constants} in the
external format to have multiple representations if they should not affect the
semantics of the data. For example, in a text protocol, the amount of
white-space separating tokens should not affect the meaning of the data and
consequently Nail does not expose it to the application. As long as
constants are only used for their intended purpose of representing syntax-only
features, the generated output will have the same semantics as the parsed input.

Third, Nail provides support for structural dependencies, transparently handling
offset and length fields. By hiding the offset and length values from the
programmer, Nail ensures they remain consistent even if the data is changed by
the application.

We have implemented an initial prototype of Nail for C\@.  Our experience
so far suggests that it is a promising approach: we were able to construct
a succinct grammar for DNS packets, and write a small DNS server that
uses Nail for all packet input and output, and operates purely on
Nail-generated data structures, with no need for any semantic actions.

The rest of this paper is organized as follows.  \S\ref{s:relwk}
puts Nail in the context of related work.  \S\ref{s:design} describes
Nail's design.  \S\ref{s:impl} discusses our initial implementation
of Nail.  \S\ref{s:eval} provides some initial evaluation results.
\S\ref{s:future} suggests several directions for future work, and
\S\ref{s:concl} concludes.

>>>>

Code that handles untrusted inputs, such as processing network
data or parsing a file, is error-prone and is often exploited by
attackers.  This is in part because attackers have precise control
over the inputs to that code, and can craft inputs that trigger
subtle corner cases in input processing.  For example, the libpng image decompression
library has had 24 remotely exploitable vulnerabilities from 2007 to
2013,\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-7294/Libpng.html}}
and Adobe's PDF and Flash viewers have been notoriously
plagued by input processing vulnerabilities.  Even relatively
simple formats, such as those used by the zlib compression
library, have had input processing vulnerabilities in the
past.\footnote{\url{http://www.cvedetails.com/vulnerability-list/vendor_id-72/product_id-1820/GNU-Zlib.html}}

@@

Third, data formats, for example PNG or PDF, contain features that are hard to represent with
existing parser generators. Some features, such as length and checksum fields, require the parser to
keep additional state. More complicated formats incorporate offset fields that require the parser to
reposition its input stream or consume the same input multiple times. In some cases, such as with
compressed data, very complicated transformations on the external format are necessary before it can
be understood. State-of-the-Art parser generators provide at most limited, unelegant facilities for
such features, so programmers are forced to manually writing control code to handle them, such as
moving the parsers input stream, looping over the invocation of a sub-parser or computing a checksum
over raw input bytes. Besides leaving much room for errors with offset arithmetic, such code is
usually not reusable when generating output.

\XXX[Motivate Stream transforms better]

@@

Second, this well-defined internal representation allows Nail to invert the
parser and generate output, while maintaining a \emph{semantic bijection}.
A bijection in the traditional sense often does not make sense when parsing data formats. 
Consider a grammar for a
text language that tolerates white space, or a binary protocol that tolerates
arbitrarily long padding.\footnote{Say, the physical layer of most communication
  protocols is a possibly infinite sequence of symbols that are syntactically
  nil followed by a pre-determined synchronization sequence and the actual
  contents of the transmission.} Program semantics should be
independent of the number of padding elements in the input, and Nail therefore
does not expose that information to the programmer. We call such discarded
fields \emph{constants}. 
Similarly, Nail does not preserve the layout of
objects referred to by their offsets. If the grammar consists of a simple protocol without offset
fields, constants and the like, there is a normal bijection between the internal model and valid
parser inputs.

\XXX[Currently, Nail always makes a default choice when
there are multiple options to express a constant, however Nail could be extended
to allow a grammar-specific plug-in to make these choices, say for faster
alignment, streaming applications when data is not ready, or visual appearance in
human-readable protocols.]

 However, Nail allows \emph{constants} in the
external format to have multiple representations if they should not affect the
semantics of the data. For example, in a text protocol, the amount of
white-space separating tokens should not affect the meaning of the data and
consequently Nail does not expose it to the application. As long as
constants are only used for their intended purpose of representing syntax-only
features, the generated output will have the same semantics as the parsed input.

Third, Nail introduces two abstractions, \emph{dependent fields} and \emph{stream transforms}, to
elegantly handle problematic features of data formats. Dependent fields are fields in the protocol
whose value depends in some way on the value or layout of other parts of the protocol, such as
length, offset and checksum fields. \XXX[Stream transforms allow the programmer to write a function
that produces one or more virtual streams from one or more virtual streams and the values of
dependent fields.]

We have implemented a prototype of Nail for C\@.  Our experience
so far suggests that it is a promising approach: we were able to construct
a succinct grammar for a set of common file formats and protocols that demonstrate Nail's ability to
handle tricky constructs and build a series of sample applications with these grammars.

The rest of this paper is organized as follows.  \S\ref{s:relwk}
puts Nail in the context of related work.  \S\ref{s:design} describes
Nail's design.  \S\ref{s:impl} discusses our initial implementation
of Nail.  \S\ref{s:eval} provides evaluation results.
\S\ref{s:future} suggests several directions for future work, and
\S\ref{s:concl} concludes.
